---
title: "p8105_hw3_yl4358"
author: "YI LIU"
date: "10/6/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

# Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```

* The `instacart` dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns; Key variables include `user_id`, `product_name`, `order_number`, `order_dow`, `order_hour_of_day` and so on. Every observation illustrates the product name, its order time, user's relevant ordering information, its belonging aisle and department.

```{r}
instacart_1 =
  instacart %>%
  group_by(aisle) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

head(instacart_1,10) %>% 
knitr::kable()
```

* There are `r nrow(instacart_1)` aisles there and `fresh vegetables` and `fresh fruits` aisles are the most items ordered from.

```{r}
instacart_2 =
  instacart %>%
  group_by(aisle) %>% 
  summarize(n = n()) %>%
  filter(n >= 10000) %>% 
   ggplot(aes(x = aisle, y = n)) + 
    geom_bar(stat = "identity", fill = "coral") + 
  labs(
    title = "The number of items ordered in each aisle",
    x = "Aisle ",
    y = "The number of items ordered"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5))
instacart_2
```

* We use bar chart to display the number of items ordered in each aisle, they are quite different among aisles. As we can see, fresh fruits aisle and fresh vegetables aisle are aisles most items ordered from. Plot is so clear and straightforward that we could get information from it.

```{r}
instacart_3 =
  instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>% 
  filter(min_rank(desc(n)) == 1) %>% 
  rename(most_popular_items = product_name, ordering_counts = n) %>% 
  knitr::kable()
instacart_3
```

* Concluded from the table, most popular items in `baking ingredient`, `dog food care` and `packaged vegetables fruits` aisle are respectively `light brown sugar`, `snack sticks chicken & rice recipe dog treats` and `organic baby spinach`.

```{r}
instacart_4 = 
  instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_hour_of_the_day = round(mean(order_hour_of_day), digits = 1)
  ) %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour_of_the_day"
  ) %>%
  rename( Monday = '1', Tuesday = '2', Wednesday = '3', Thursday = '4', Friday = '5', Saturday = '6', Sunday = '0') %>% 
  knitr::kable()

instacart_4
```

* People are more likely to order `Coffee Ice Cream` during 12:00pm-15:00pm which is teatime and order `Pink Lady Apples` during 11:00am to 14:00 pm when having lunch.


# Problem 2

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_smart2010 = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health" ) %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>% 
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))
  )
```

```{r}
states_grt7_2002 =
  brfss_smart2010 %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarize(
    n_unique = n_distinct(locationdesc)
  ) %>% 
  filter(n_unique >= 7) %>% 
  rename(state = locationabbr, number_of_locations = n_unique) %>% 
  knitr::kable()

states_grt7_2002

states_grt7_2010 =
  brfss_smart2010 %>% 
  filter(year == 2010) %>% 
  group_by(locationabbr) %>% 
  summarize(
    n_unique = n_distinct(locationdesc)
  ) %>% 
  filter(n_unique >= 7) %>% 
  rename(state = locationabbr, number_of_locations = n_unique) %>% 
  knitr::kable()

states_grt7_2010
```

* In 2002, `CT`, `FL`, `MA`, `NC`, `NJ` and `PA` were observed at 7 or more locations; In 2010, `CA`, `CO`, `FL`, `MA`, `MD`, `NE`, `NJ`, `NY`, `OH`, `PA`, `SC`, `TX`, `WA` and `NC` were observed at 7 or more locations.


```{r}
excellent_df = 
  brfss_smart2010 %>% 
  filter(response == "Excellent") %>% 
  select(year, locationabbr, locationdesc, data_value) %>% 
  group_by(year, locationabbr) %>% 
  summarize(
    average_value = mean(data_value)
  ) %>% 
  rename(state = locationabbr)

excellent_df %>% 
  ungroup() %>% 
  group_by(state) %>% 
  ggplot(aes(x = year, y = average_value, color = state)) + 
  geom_line() +
  labs(
    title = "Average value over time within a state",
    x = "Year",
    y = "Average data value"
  ) +
  theme_bw() 
  
```

* The average data values vary greatly among different states.

```{r}
brfss_smart2010 %>% 
  filter(year == 2006 | year == 2010) %>%
  filter(locationabbr == "NY") %>% 
  group_by(year, locationdesc) %>%
  ggplot(aes(x = response, y = data_value, color = locationdesc, group = locationdesc)) +
  geom_point() + geom_line()+
  labs(
    title = "Distribution of data value for response among locations in NY States",
    x = "Response",
    y = "Data value"
  ) +
  theme_bw() +
  theme(legend.position = "bottom") +
  facet_grid(.~year)
  
```

* Concluded from the plot, `Good`, `Very good` and `Excellent` responses have high data value while `Poor` and `Fair` responses have low data value among locations in NY state.


# Problem 3

```{r}
accel_data = read_csv(file = "./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    day_category = case_when(
      day == "Monday" ~ "weekday",
      day == "Tuesday" ~ "weekday",
      day == "Wednesday" ~ "weekday",
      day == "Thursday" ~ "weekday",
      day == "Friday" ~ "weekday",
      day == "Saturday" ~ "weekend",
      day == "Sunday" ~ "weekend",
      TRUE     ~ ""
  )
  ) %>% 
  select(week, day_id, day, day_category, everything()) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_id", 
    names_prefix = "activity_",
    values_to = "activity_counts"
  ) %>% 
  mutate(
  activity_counts = round(activity_counts, digit = 0),
  activity_counts = as.integer(activity_counts)
) 
```

* `accel_data` dataset has `r nrow(accel_data)` rows and `r ncol(accel_data)` columns. The resulting dataset has `r names(accel_data)` variables and `r length(accel_data$week)` observations.

```{r}
total_activity_accel = 
  accel_data %>% 
  group_by(week, day) %>% 
  summarize(
    total_activity_count = sum(activity_counts)
  ) 
  
 table_activity = 
   total_activity_accel %>% 
   pivot_wider(
    names_from = "day",
    values_from = "total_activity_count"
  ) %>% 
  select(week, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday) %>% 
  knitr::kable()
  
table_activity
```

* As seen from the table, trends are not apparent.

```{r}
accel_data %>%
group_by(week,day) %>% 
  mutate(
    activity_id = as.numeric(activity_id)
  ) %>% 
  ggplot(aes(x = activity_id, y = activity_counts, color = day)) +
  geom_area() + 
  labs(
    title = "24-hour activity for each day",
    x = "24-hour activity",
    y = "Activity counts"
  ) +
  viridis::scale_color_viridis(
    name = "Day", 
    discrete = TRUE
  ) +
   theme_bw() +
  theme(legend.position = "bottom") +
  facet_grid(.~week)
```

* Inspection of activity of this 63 year-old male is more steady from Tuesday to Thursday, while changes greatly from Friday to Monday.
